---
title: "Replication of Figure I–J: YAP1, TAZ, and TEAD4 Co-occupancy at Promoters and Enhancers"
author: "Harshini Muthukumar"
date: "2026-03-19"
output: html_document
---


### Figure 1I

From the figure description:

**Heatmap representing YAP/TAZ/TEAD binding sites located on promoters (top) and enhancers (bottom)**. YAP, TAZ and TEAD4 peaks are ranked from the strongest to weakest signal in TAZ ChIP, in a window of ±1kb centered on the summit of TAZ peaks. H3K4me1 and H3K4me3 signal in the corresponding genomic regions is shown on the right.

From the method section:

Heatmaps were generated using a custom R script which considers a **2-kb window centered on peak summits and calculates the normalized reads density with a resolution of 50 bp**.

The rows are different peaks. The columns are 2kb window centered on the summit of the TAZ peaks with 50 bp per bin. So it will be 2000/50 = 40 bins for the columns.The signal is the normalized read counts in each bin. 



```{r,load_peaks,echo=TRUE}
suppressWarnings(
  suppressMessages({
    library(rtracklayer)
    library(here)
    library(dplyr)
    library(ggplot2)
    library(EnrichedHeatmap)
  })
)

TAZ_peaks<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TAZ_peaks.narrowPeak"))
YAP_peaks<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/YAP_peaks.narrowPeak"))
TEAD4_peak<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TEAD4_peaks.narrowPeak"))

YAP_overlap_TAZ_peaks<-suppressWarnings(subsetByOverlaps(YAP_peaks, TAZ_peaks))

YAP_overlap_TAZ_peaks_overlap_TEAD4<- suppressWarnings(subsetByOverlaps(YAP_overlap_TAZ_peaks, TEAD4_peak))



```
Since I focused on YAP1, I will use the summit of YAP1 as the anchor point (can of course change it to TAZ too)


```{r,yap_summit,echo=TRUE}
YAP_summit<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/YAP_summits.bed"))
```

```{r,histone_markers,echo=FALSE}
H3K4me1<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/H3K4me1.bed"))
H3K4me3<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/H3K4me3.bed"))
H3K27ac<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/H3K27ac.bed"))
```

```{r,define_regions,echo=FALSE}
enhancers<- subsetByOverlaps(H3K4me1, H3K4me3, invert=TRUE)

promoters<- subsetByOverlaps(H3K4me3, H3K4me1, invert=TRUE)
```

```{r,yap_regions,echo=FALSE}
YAP1_enhancers<- suppressWarnings(subsetByOverlaps(YAP_overlap_TAZ_peaks_overlap_TEAD4, enhancers))

 
YAP1_promoters <- suppressWarnings(subsetByOverlaps(YAP_overlap_TAZ_peaks_overlap_TEAD4, promoters))

YAP1_enhancers$name %>% head()
```
```{r,combine_yap_summit,echo=FALSE}
YAP_summit_enhancer<- YAP_summit[YAP_summit$name %in% YAP1_enhancers$name]
YAP_summit_promoter<- YAP_summit[YAP_summit$name %in% YAP1_promoters$name]

# combine them
anchors<- c(YAP_summit_promoter, YAP_summit_enhancer) 
```



```{r,load_bw,echo=FALSE}
YAP1_bw<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/YAP1.bw"))
TAZ_bw<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TAZ.bw"))
TEAD4_bw<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TEAD4.bw"))

```
```{r,normalize,echo=FALSE}
#BiocManager::install("EnrichedHeatmap")
# extend 1000 bp on each side and use 50bp bin
mat1<- normalizeToMatrix(YAP1_bw, anchors, value_column = "score",
                         extend= 1000, mean_mode = "w0", w=50)

mat2<- normalizeToMatrix(TAZ_bw, anchors, value_column = "score",
                         extend= 1000, mean_mode = "w0", w=50)

mat3<- normalizeToMatrix(TEAD4_bw, anchors, value_column = "score",
                         extend= 1000, mean_mode = "w0", w=50)
```

```{r,mat,echo=TRUE}
dim(mat1)
dim(mat2)
dim(mat3)
```
```{r,echo=TRUE}
mat1[1:5, 1:40 ]
```

```{r,quantile,echo=TRUE}
quantile(mat1, c(0.1,0.25,0.5,0.9,1))
quantile(mat2, c(0.1,0.25,0.5,0.9,1))
quantile(mat3, c(0.1,0.25,0.5,0.9,1))
```

```{r,color,echo=FALSE}
col_fun<- circlize::colorRamp2(c(0, 20), c("white", "red"))
```

```{r,partition,echo=TRUE}
partition<- c(rep("promoters", length(YAP1_promoters)),
              rep("enhancers", length(YAP1_enhancers)))

# change the factor level so promoters come first
partition<- factor(partition, levels=c("promoters", "enhancers"))

partition_hp<- Heatmap(partition, col=structure(2:3, names = c("promoters", "enhancers")), 
        name = "partition",
        show_row_names = FALSE, width=unit(3,'mm'))

partition_hp
```
```{r,heatmap,echo=TRUE}
ht_opt$message = FALSE

ht_list<- partition_hp +
  EnrichedHeatmap(mat1, pos_line = FALSE, column_title="YAP1", name = "YAP1", col=col_fun) +
  EnrichedHeatmap(mat2, pos_line = FALSE, column_title="TAZ", name = "TAZ", col=col_fun) +
  EnrichedHeatmap(mat3, pos_line = FALSE, column_title="TEAD4", name = "TEAD4", col=col_fun)

draw(ht_list, split= partition, main_heatmap =2)
```


The heatmaps depict ChIP-seq binding enrichment of YAP1, TAZ, and TEAD4 across promoter and enhancer regions, centered at the start site spanning ±1000 bp, with red intensity reflecting signal strength. All three factors show strong, sharp co-occupancy at promoters, as evidenced by the tight peaks in the profile plots, with TEAD4 displaying the most defined and focused binding signal among the three. At enhancers, binding is present across all factors but notably more diffuse and variable across regions. The broad co-occupancy of these factors is consistent with the known biology of the Hippo signaling pathway, wherein YAP1 and TAZ function as transcriptional co-activators that work through TEAD family members such as TEAD4 to regulate target gene expression at both promoters and distal regulatory elements.

## FIG 1 J

Figure description:

Bimodal distribution of H3K4me1 signal around the summit of YAP/TAZ and TEAD4 peaks.

The line plot is just the average of all the peaks per column.

```{r,mean,echo=TRUE}
YAP1_mean<- colMeans(mat1)
TAZ_mean<- colMeans(mat2)
TEAD4_mean<- colMeans(mat3)
```

```{r,plot,echo=TRUE}
plot(YAP1_mean)
```

```{r,echo=TRUE}
bind_rows(YAP1_mean, TAZ_mean, TEAD4_mean) %>%
  mutate(factor = c("YAP1", "TAZ", "TEAD4")) %>%
  select(factor, everything())
```
```{r,bind_Row,echo=TRUE}
bind_rows(YAP1_mean, TAZ_mean, TEAD4_mean) %>%
  mutate(factor = c("YAP1", "TAZ", "TEAD4")) %>%
  select(factor, everything()) %>%
  tidyr::pivot_longer(-factor)
```
```{r,line_plot,echo=TRUE}
bind_rows(YAP1_mean, TAZ_mean, TEAD4_mean) %>%
  mutate(factor = c("YAP1", "TAZ", "TEAD4")) %>%
  select(factor, everything()) %>%
  tidyr::pivot_longer(-factor) %>%
  ggplot(aes(x=name, y=value)) +
  geom_line(aes(color = factor, group=factor))
```
```{r,factor,echo=TRUE}
bind_rows(YAP1_mean, TAZ_mean, TEAD4_mean) %>%
  mutate(factor = factor(c("YAP1", "TAZ", "TEAD4"), levels = c("YAP1", "TAZ", "TEAD4"))) %>%
  select(factor, everything()) %>%
  tidyr::pivot_longer(-factor) %>%
  mutate(name = factor(name, levels = c(paste0("u",1:20), paste0("d", 1:20)))) %>%
  ggplot(aes(x=name, y=value)) +
  geom_line(aes(color = factor, group=factor)) +
  scale_x_discrete(breaks=c("u1", "d1", "d20"), labels = c("-1kb", "0", "1kb")) +
  scale_color_manual(values = c("#DA9195", "#E07B78", "#605D7D")) +
  theme_classic(base_size = 14) +
  ylab("RPKM") +
  xlab("")
```

