---
title: "Replication of Figure A,B,C:Common Peaks Between TAZ, YAP, and TEAD4: Genomic Distribution and Signal Concordance"
author: "Harshini Muthukumar"
date: "2026-03-19"
output: "html_document"
---

### Figure 1a

Figure 1a and 1b are venn-diagram showing the overlapping peak number of YAP/TAZ and YAP/TAZ/TEAD4

To find overlap/intersection of genomic regions, one can use the famous bedtools on command line or use GenomicRanges in R.

```{r,peak_files,echo=FALSE}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(rtracklayer)
  library(here)
  library(VennDiagram)
  library(grid)
  library(dplyr)
  library(ggplot2)
})


TAZ_peaks<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TAZ_peaks.narrowPeak"))
YAP_peaks<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/YAP_peaks.narrowPeak"))


```


```{r,taz_peak,echo=TRUE,warning=FALSE, message=FALSE}
TAZ_overlap_YAP_peaks<- subsetByOverlaps(TAZ_peaks, YAP_peaks)
length(TAZ_overlap_YAP_peaks)
```


```{r,yap_taz,echo=TRUE,warning=FALSE, message=FALSE}
YAP_overlap_TAZ_peaks<- subsetByOverlaps(YAP_peaks, TAZ_peaks)
length(YAP_overlap_TAZ_peaks)
```

```{r,overlap,echo=TRUE}

n_YAP <- length(YAP_peaks)
n_TAZ <- length(TAZ_peaks)
n_overlap <- length(YAP_overlap_TAZ_peaks)

grid.newpage()

venn.plot <- draw.pairwise.venn(
  area1 = n_YAP,
  area2 = n_TAZ,
  cross.area = n_overlap,
  category = c("YAP", "TAZ"),
  fill = c("skyblue", "pink"),
  alpha = 0.5,
  cex = 2,
  cat.cex = 2
)

grid.draw(venn.plot)
```

## Figure 1: Overlap between TAZ and YAP ChIP-seq peaks

YAP (Yes-Associated Protein) and TAZ (Transcriptional Co-Activator with PDZ-binding motif) are the two core effectors of the Hippo signaling pathway. They are paralogs — meaning they share significant structural similarity and are often assumed to bind the same genomic regions.

However, whether they truly co-occupy the same sites across the genome, or have distinct.binding preferences, is a key biological question.To address this, we called peaks from ChIP-seq experiments for both YAP and TAZ separately and then asked: how much do their binding sites overlap? The Venn diagram shows that out of all peaks identified, **7,164 are shared** between YAP and TAZ, while **3,555 are TAZ-specific** and **2,643 are YAP-specific**. 

The majority of binding sites (~67%) are co-occupied, consistent with their role as functional paralogs that often work together. However, the unique sites suggest each protein may also have independent regulatory roles — TAZ appears 
to have more unique binding sites than YAP in this dataset, hinting at divergent functions beyond their shared targets.

```{r,TEAD4_overlap,echo=TRUE}
TEAD4_peak<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TEAD4_peaks.narrowPeak"))

YAP_overlap_TAZ_peaks_overlap_TEAD4<- subsetByOverlaps(YAP_overlap_TAZ_peaks, TEAD4_peak)


n_YAP_TAZ <- length(YAP_overlap_TAZ_peaks)
n_TEAD4 <- length(TEAD4_peak)
n_overlap2 <- length(YAP_overlap_TAZ_peaks_overlap_TEAD4)

grid.newpage()

venn.plot2 <- draw.pairwise.venn(
  area1 = n_YAP_TAZ,
  area2 = n_TEAD4,
  cross.area = n_overlap2,
  category = c("YAP/TAZ", "TEAD4"),
  fill = c("lightgreen", "orange"),
  alpha = 0.5,
  cex = 2,
  cat.cex = 2
)

grid.draw(venn.plot2)


```

## Figure 2: Overlap between TEAD4 and YAP/TAZ binding sites

YAP and TAZ are **transcriptional co-activators** — they do not bind DNA directly. Instead,they work by physically interacting with DNA-binding transcription factors, most prominently the TEAD family (TEAD1–4). TEAD4 is the dominant TEAD family member in many cell types and serves as the primary DNA anchor that recruits YAP/TAZ to target gene promoters and enhancers.

To validate that the YAP/TAZ peaks we identified are functionally meaningful, we asked: do the combined YAP/TAZ binding sites (from Figure 1) overlap with regions bound by TEAD4? The Venn diagram shows that **5,965 sites are shared** between TEAD4 and YAP/TAZ, representing the canonical YAP/TAZ–TEAD co-bound regions that are most likely to be actively regulating transcription. 

TEAD4 has **5,547 unique sites** — these may reflect TEAD4 binding with other co-factors independently of YAP/TAZ. Only **1,199 YAP/TAZ sites** lack TEAD4, which could represent TEAD-independent YAP/TAZ interactions (e.g., via AP-1 or other partners). This overlap analysis gives us confidence that our ChIP-seq peaks are capturing biologically relevant co-regulatory complexes.

### Fig 1C

```{r,summit,echo=FALSE}
TAZ_summit<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TAZ_summits.bed"))

TAZ_summit<- TAZ_summit[TAZ_summit$name %in% TAZ_overlap_YAP_peaks$name]

TEAD4_summit<- import(here("C:/Users/harzi/OneDrive/Desktop/chip_seq_peak/TEAD4_summits.bed"))

```
```{r,hits,echo=FALSE}
TAZ_500bp_window<- resize(TAZ_summit, width = 500, fix="center")

hits<- findOverlaps(TEAD4_summit, TAZ_500bp_window)

```
```{r,dist,echo=FALSE}
summit_distance<- distance(TEAD4_summit[queryHits(hits)], TAZ_summit[subjectHits(hits)])

```
```{r,tead4_summit,echo=TRUE}
TEAD4_summit[queryHits(hits)][summit_distance ==0]
```

```{r,taz_summit,echo=TRUE}
TAZ_summit[subjectHits(hits)][summit_distance ==0]
```

```{r,signed_dist,echo=FALSE}
# Compute signed distances
signed_distance <- function(A, B) {
  # Compute unsigned distance
  dist <- distance(A, B)
  
  # Determine signs based on whether A precedes or follows B
  sign <- ifelse(start(A) < start(B), -1, 1)
  
  # Apply sign to distance
  dist * sign
}

```

```{r,dist_df,echo=TRUE}

summit_distance<- signed_distance(TEAD4_summit[queryHits(hits)],
                                  TAZ_summit[subjectHits(hits)])

distance_df<- table(summit_distance) %>%
  tibble::as_tibble() 

distance_df
```
```{r,geom_plot,echo=TRUE}
distance_df %>%
  ggplot(aes(x=summit_distance, y = n)) +
  geom_point()
```

```{r,plot,echo=TRUE}
distance_df %>%
  mutate(summit_distance = as.numeric(summit_distance)) %>%
  arrange(summit_distance) %>%
  ggplot(aes(x=summit_distance, y = n)) +
  geom_point()
```
```{r,plot_2,echo=TRUE}
distance_df %>%
  mutate(summit_distance = as.numeric(summit_distance)) %>%
  arrange(summit_distance) %>%
  ggplot(aes(x=summit_distance, y = n)) +
  geom_line()
```

```{r,df_binned,echo=FALSE}
df_binned <- distance_df %>%
  mutate(summit_distance = as.numeric(summit_distance)) %>%
  arrange(summit_distance) %>%
  mutate(bin = floor(summit_distance / 5) * 5) %>%  # Create bins by grouping every 5 bp
  group_by(bin) %>%
  summarise(n = mean(n, na.rm = TRUE))  # Calculate average 'n' for each bin
```

```{r,plot_3,echo=FALSE}
df_binned %>%
  ggplot(aes(x=bin, y = n)) +
  geom_line() +
  scale_x_continuous(breaks = c(-250, 0, 250)) +
  xlab("distance to the summit \nof TAZ peaks (bp)") +
  ylab("peak density") +
  theme_classic(base_size = 14) 
```


## Figure 3: Peak density relative to TAZ summit

This plot shows how ChIP-seq signal is distributed around the center (summit) 
of TAZ binding sites within a ±250 bp window. 

The sharp spike at position 0 tells us that the signal is tightly concentrated 
exactly at the binding site — this is what we expect for a transcription factor 
like TAZ, which binds short, specific DNA sequences. The quick drop-off on both 
sides confirms the data is clean and the immunoprecipitation worked well. 
This figure serves as both a **quality check** for the ChIP-seq experiment and 
evidence that TAZ is binding to precise, well-defined locations in the genome.

